---
title: Semantic fluency data - Overview
output: pdf_document
---

```{r global_options, include=FALSE}
    knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE, dpi=300)
    library(data.table)
    library(ggplot2)
    library(gridExtra)
    library(knitr)

    dat <- fread('../results_cleaned.csv')
    dat <- dat[shortlist==0]            # exclude categories that had to be re-done for now
```

These first few graphs are a cleaned version of previous data. Plurals or spelling mistakes
by the same person by have been merged, resulting in many fewer unique items.

- ### `r length(dat[,item])` items were listed in total, consisting of `r length(dat[,unique(item)])` unique items.

## Total items by category

```{r}
    kable(dat[,.(N=.N,"Unique N"=length(unique(item))),by=category])
```

## Average number of unique items listed per category

```{r}
    kable(dat[,length(unique(item)) ,by=.(id,category)][,.(mean=mean(V1),median=median(V1)),by=category])
```

## Number of items listed per round by category

```{r}
    #dat <- dat[!(id=="S101" & game==0)] # exclude RT of S101 game 0; he entered a bunch of items on the same line
    kable(dat[,.N,by=.(category,game,id)][,.(mean=mean(N),median=as.double(median(N))),by=category])
```

# Number of animals listed as a function of time into round
```{r}
ggplot(data=dat[category=="animals"],aes(x=RTstart/1000)) + geom_histogram(binwidth=5) +
    xlab("Seconds into round") + ylab("Count")
```

# Distribution of animal response times
```{r}
ggplot(data=dat[category=="animals"],aes(x=RT/1000)) + geom_histogram(binwidth=0.5) +
    xlab("RT in seconds") + ylab("Count") + scale_x_continuous(limits=c(0,20))
```

# How many animals were listed once, twice, etc. across people
```{r}
    plotdat <- dat[,.(.N,category),by=.(id,item)][,.N,by=.(item,category)]
    ggplot(data=plotdat[category=="animals"],aes(x=N)) + geom_histogram(binwidth=1) +
        xlab("Number of times animal was listed") + ylab("Count")
```

# Distribution of number of unique items listed by category
```{r}
    animals_plot <- dat[category=="animals",length(unique(item)) ,by=.(id,category)]
    vegetables_plot <- dat[category=="vegetables",length(unique(item)) ,by=.(id,category)]
    fruits_plot <- dat[category=="fruits",length(unique(item)) ,by=.(id,category)]
    p1 <- ggplot(dat=animals_plot,aes(x=V1)) + geom_histogram(position="dodge",binwidth=10) + xlab("Number of unique animals listed") + ylab("People")
    p2 <- ggplot(dat=vegetables_plot,aes(x=V1)) + geom_histogram(position="dodge",binwidth=10) + xlab("Number of unique vegetables listed") + ylab("People")
    p3 <- ggplot(dat=fruits_plot,aes(x=V1)) + geom_histogram(position="dodge",binwidth=10) + xlab("Number of unique fruits listed") + ylab("People")
    grid.arrange(p1,p2,p3,nrow=3)   
```

# Proportion of animal semantic space (of each user) covered by a single X
```{r,echo=FALSE}
    a<-dat[,.(uniquetotal=length(unique(item))),by=.(category,id)]
    b<-dat[,.(uniquepergame=length(unique(item))),by=.(category,id,game)]
    setkey(b,id,category,game)
    setkey(a,id,category)
    a<-a[b]
    invisible(a[,propX:=(uniquepergame/uniquetotal)])
    invisible(a[,meanpropXbyid:=mean(propX),by=id][order(meanpropXbyid)])
    p4 <- ggplot(dat=a[category=="animals"],aes(x=meanpropXbyid)) + geom_histogram(binwidth=.05) + 
          xlab("Proportion of semantic space covered") + ylab("Number of Xs")
    print(p4)
```

# Average proportion of space covered by category
```{r}
    a[,meanpropXsubj:=mean(propX),by=.(id,category)][,mean(meanpropXsubj),by=category]
```

